---
title: "Rockfish_Results"
format: pdf
editor: visual
---

## Goals of this study

Investigate patterns in the acoustic behavior of rockfish

## Data

Acoustic data was collected from stationary recorders along the central California coast. Site = MB05

## Combine Logs

The presence of rockfish chorusing was logged by 2 Seatech analysts: Rie and Gale.

```{r}
#| echo: false
#| message: false
#| warning: false

# # # if (!require("pacman")) install.packages("pacman")
if(!require('devtools')) install.packages('devtools')
# # install from GitHub
devtools::install_github('TaikiSan21/PAMmisc')
# devtools::install_github('ericpante/marmap')
suppressMessages(pacman::p_load(knitr,here,lubridate,dplyr,magrittr,xlsx,RSQLite,viridis,marmap,ggplot2,suncalc,ggnewscale,PAMmisc,stringr,lunar,NightDay))

Logs<-list.files(here('data','2024_Rockfish'),pattern='detlog',full.names = TRUE) %>%
  lapply(read.xlsx,sheetName='Detections') %>%
  bind_rows() %>%
  mutate(StartTime=as.POSIXct(Start.time,format='%Y-%m-%d %H:%M:%S',tz='UTC'),
         EndTime=as.POSIXct(End.time,format='%Y-%m-%d %H:%M:%S',tz='UTC'))

# Logs<-Logs %>%
#   mutate(interval=floor_date(PT,unit="hour")+minutes(floor(minute(PT)/10)*10), Hour=hour(PT)) 

#Create Times Dataframe
EffortStart<-as.POSIXct('2023-03-01 0:00:00',format='%Y-%m-%d %H:%M:%S',tz='UTC')
EffortEnd<-as.POSIXct('2023-03-31 23:59:59',format='%Y-%m-%d %H:%M:%S',tz='UTC')
Times<- seq.POSIXt(EffortStart,EffortEnd,by='10 min')
MB05<-data.frame(Times)
MB05$Rockfish<-'Absent'

for(n in 1:nrow(Logs)){
  pres<-which(MB05$Times>=Logs$Start.time[n] & MB05$Times<Logs$End.time[n])
  MB05$Rockfish[pres]<-'Present'
}

MB05$LunarDay<-lunar.phase(MB05$Times,name=8)

#Add Day/Night values
MB05$DayNight<-'night'
SunTimes<-getSunlightTimes(unique(as.Date(MB05$Times)),lat=35.883333,
                lon=-121.55,keep=c('sunrise','sunset'),tz='America/Los_Angeles')

for(n in 1:length(SunTimes)){
  day<-which(MB05$Times>=SunTimes$sunrise[n] & MB05$Times<MB05$sunset[n])
  MB05$DayNight[day]<-'day'
}
MB05$Times_local<-with_tz(MB05$Times,tz='America/Los_Angeles')
```

## Mark times of no-effort due to masking

At times, noise masked the presence of rockfish chorusing so we need to show that these times have 'no effort'.

```{r}
#| echo: false
#| message: false
#| warning: false

#Start and end of noise 
NoiseStart<-with_tz(as.POSIXct('2023-03-21 06:23',tz='UTC'),tz='America/Los_Angeles')
NoiseStop<-with_tz(as.POSIXct('2023-03-27 14:00',tz='UTC'),tz='America/Los_Angeles')
```

## Plots

**Total call hours by date with Lunar Illumination**

```{r}
Logs<-Logs%>%
  mutate(Duration = difftime(End.time,Start.time,units="hour"),
         LunarDay =  round(lunar.phase(as.Date(StartTime))/0.212769,0))
plotfile<-paste0(here('figures','Rockfish_Chorus_Duration_Lunar.pdf'))

pdf(plotfile,width=6,height=6)           
ggplot(Logs,aes(LunarDay,Duration))+
  geom_point()+xlab('Lunar Day')+theme_bw()+geom_smooth()+
  ggtitle('Lunar Cycle with Chorus Duration')
dev.off()

Summary<-MB05 %>% 
  filter(Rockfish=="Present") %>%
  mutate(Date = as.Date(Times_local)) %>%
  mutate(LunarDay =  round(lunar.phase(as.Date(Times_local))/0.212769,0),
         LunarIll = lunar.illumination(Date))%>%
  group_by(Date) %>%
  summarize(HourPres = n()/6,
            LunarDay = median(LunarDay),
            LunarIll = mean(LunarIll)) 

plotfile<-paste0(here('figures','Rockfish_lunar.pdf'))

pdf(plotfile,width=6,height=4.5)
print(ggplot(Summary,aes(Date,HourPres,fill=LunarIll))+
  annotate("rect", fill = "grey",  
        xmin = as.Date(NoiseStart), xmax = as.Date(NoiseStop),
        ymin = 0, ymax = Inf) +
  geom_bar(stat='identity',color='black') +
  scale_fill_gradient2(low='black',high='white',midpoint=.5)+
  ylab('Total hours with calling')+labs(fill='Lunar Illumination'))
dev.off()

ggplot(Summary,aes(LunarDay,HourPres,fill=LunarIll))+
  geom_bar(stat='identity',color='black') +
  scale_fill_gradient2(low='black',high='white',midpoint=.5)
```

#### Diel plot showing rockfish presence

```{r}
#| echo: false
#| message: false
#| warning: false
GPS<-data.frame(Latitude=rep(35.7668,2),Longitude=rep(-121.4332,2),UTC=c(as.POSIXct('2023-03-01 0:00',tz='UTC'),as.POSIXct('2023-03-31 23:59',tz='UTC')),DriftName=rep('MB05',2))

MB05Pres<-MB05 %>%
  filter(Rockfish=='Present') %>%
  mutate(UTC=Times)

plotPresGrid(MB05Pres,type="presence",bin="15min",gps=GPS,title=paste0('MB05: ',': Rockfish'),plotTz='America/Los_Angeles') 

MB05Pres <-MB05Pres%>%
    mutate(lunar_day = round(lunar.phase(as.Date(Times)) / 0.212769),0) 

```

### Time with soundscape metrics

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 8

library(lubridate)
library(tidyverse)
library(grid)
library(gridExtra)
library(scales)
library(data.table)
library(reshape)
library(viridis)
SSDir<-'D:/Seatech/Soundscape/metrics/'
TOLfilelist<-list.files(path=SSDir,pattern='*TOL_2min',recursive=TRUE,full.names = TRUE)
TOL<-lapply(TOLfilelist,read.csv,header=TRUE)
TOL<-bind_rows(TOL)
TOL$DateTime<-strptime(TOL$yyyy.mm.ddTHH.MM.SSZ,format="%Y-%m-%dT%H:%M:%OSZ",tz="UTC")
TOL$DateTime<-with_tz(TOL$DateTime,tz='America/Los_Angeles')
TOL$DateTime<-as.POSIXct(TOL$Date)
TOL<-select(TOL, -yyyy.mm.ddTHH.MM.SSZ)
TOL<-melt(TOL,id.vars=c('DateTime'))
TOL$Month<-month(TOL$DateTime)
TOL$Date<-as.Date(TOL$DateTime,format='%Y-%m-%d')

# TOL<-read.csv(fileTOL,header = TRUE)
# TOL$DateTime<-strptime(TOL$yyyy.mm.ddTHH.MM.SSZ,format="%Y-%m-%dT%H:%M:%OSZ",tz="UTC")
# TOL$DateTime<-as.POSIXct(TOL$Date)
# TOL<-select(TOL, -yyyy.mm.ddTHH.MM.SSZ)
# TOL<-melt(TOL,id.vars=c('DateTime'))
# TOL$Month<-month(TOL$DateTime)
# TOL$Date<-as.Date(TOL$DateTime,format='%Y-%m-%d')

TOL200<-filter(TOL,variable=="TOL_200")%>%
  mutate(Hour=hour(DateTime))
TOL250<-filter(TOL,variable=="TOL_250")%>%
  mutate(Hour=hour(DateTime))

#PLOTS
ggplot(TOLsub,aes(DateTime,value,color=variable))+geom_line()+
  scale_color_viridis(discrete=TRUE)+
  geom_vline(xintercept=Logs$Start.time,col='green')+
  geom_vline(xintercept=Logs$End.time,col='red')

Heat200<-ggplot(TOL200,aes(Hour,Date,fill=value))+geom_tile()+
  scale_fill_viridis(option="H")
Heat250<-ggplot(TOL250,aes(Hour,Date,fill=value))+geom_tile()+
  scale_fill_viridis(option="H")
  

grid.arrange(Heat200,Heat250,nrow=2)
  scale_fill_viridis(name='dB re 1 uPa')+ggtitle("MB05: Third Octave Level @ 200 Hz (median)")
```

```{r}
#| echo: false
#| message: false
#| warning: false
MB05_gps<-data.frame(Latitude=rep(35.7668,2),Longitude=rep(-121.4332,2),UTC=c(as.POSIXct('2023-03-01 0:00',tz='UTC'),as.POSIXct('2023-03-31 23:59',tz='UTC')),DriftName=rep('MB05',2))

figtitle<-''

 # define boundaries of the map
 lon1=-122;lon2=-120.4;lat1=35.1;lat2=36;
  
  #Extract bathymetry data from NOAA (saves local file to speed future performance)
  bat<-getNOAA.bathy(lon1,lon2,lat1,lat2,resolution=1,keep=TRUE)
  
  #Create color palettes
  blues<-c("royalblue4","royalblue3",
           "royalblue2","royalblue1")
  greys<-c(grey(0.8),grey(0.93),grey(0.99))
  dasbr.ptColor = "black"  # color for DASBR track and deployment points
  SPLabels<-'Humpback Song'
  SpColor<-c('yellow')
  
    # png(plotfile,width=5,height=6,units="in",res=300)
    plotfile<-paste0(here('figures','MB05_Map.pdf'))

  pdf(plotfile,width=6,height=4.5)
  #Whales
 autoplot.bathy(bat,geom=c('r'),coast=TRUE)+
    coord_cartesian(expand = 0)+
    geom_contour(aes(z=z),breaks=c(-100,-500,-1000),color="gray24")+
    scale_fill_gradientn(values = scales::rescale(c(min(bat), -120, 0, max(bat))),
                         colors = c("midnightblue", "royalblue3", "grey50", "grey80"),name="Depth (m)")+
    xlab('')+ylab('')+ggtitle(figtitle)+
      new_scale_color()+
   geom_point(data=MB05_gps,aes(Longitude,Latitude),size=4,color='orange')+
   geom_text(data=MB05_gps,aes(Longitude,Latitude,label='MB05'),size=4,vjust=2)+
   geom_point(aes(-120.85,35.3659,label='Morro Bay'),size=2)+
   geom_text(aes(-120.85,35.3659,label='Morro Bay'),size=4,hjust=-.1)
dev.off()


 }
```
